Methods for Analyzing Models of\\ Rod-Shaped Bacteria

### 2025-11-30 paper/main.tex
change name of subsection again
rewrite introduction to applications again

\subsubsection{Lineage-aware Imaging-Space Metric}
\label{subsec:image-space-metric}
\paragraph{Motivation}
The preceding algorithm is performant and effective but cannot be applied in the case when a
division event splits apart the bacterium since this would require comparing sets of unequal
cardinality.
This situation is displayed in figure~\ref{fig:mask-difference-metric} (E) where two distinct
bacterial configurations originate from the same initial state but contain varying numbers of
bacteria at the next time step.
Such a mismatch can happen, e.g. due to non-matching division times of model and data.
Thus a cell which is present in the simulation might not yet have a corresponding cell in the
experimental data at this time point yet (or vice-versa).

@@

In this section we apply the previously developed methods and model to three case-studies, each
increasing in complexity.
We set out to estimate the parameters of every configuration.

First, we fit our model to data of elastically bending bacteria confined in a microfluidic
apparatus which was previously published by Amir et. al~\cite{Amir2014,Amir2014_2}.
This scenario isolates the mechanics of the flexible rod and is void of any complex cell-cell
interactions.
It is therefore well suited as a starting point in assessing the fidelity of our approach.

Subsequently, we extend our scope by incorporating multiple bacteria with cell-cell interactions
(see subsection~\ref{section:application-2}).
Initially, we consider a time interval that does not contain division events and are thus able to
employ the direct metric (see \nameref{subsec:direct-metric}).
Finally, we apply the image-space metric (see \nameref{subsec:image-space-metric}) to a time-series
containing multiple division events, demonstrating the feasibility of our approach in such
scenarios.

>>>>

\subsubsection{Lineage-aware Image-Space Metric}
\label{subsec:image-space-metric}
\paragraph{Motivation}
The preceding algorithm is performant and effective but cannot be applied in the case when a
division event splits apart the bacterium since this would require comparing sets of unequal
cardinality.
This situation is displayed in figure~\ref{fig:mask-difference-metric} (E) where two distinct
bacterial configurations originate from the same initial state but contain varying numbers of
bacteria at the next time step.
Such a mismatch can happen, e.g. due to non-matching division times of model and data.
Thus a cell which is present in the simulation might not yet have a corresponding cell in the
experimental data at this time point yet (or vice-versa).

@@

In this section we apply the previously developed methods and model to three case-studies, each
increasing in complexity.
We estimate the parameters for every configuration.

First, we fit our model to data of elastically bending bacteria published by Amir et.
al~\cite{Amir2014,Amir2014_2}.
This scenario isolates the mechanics of the flexible rod and excludes any complex cell-cell
interactions.
It is therefore well suited as a starting point in assessing the fidelity of our approach.

Subsequently, we extend our scope by incorporating multiple bacteria with cell-cell interactions
(see \nameref{section:application-2}).
Initially, we consider a time interval that does not contain division events and enabling us to
employ the direct Position-space metric (Section~\ref{subsec:direct-metric}).
Finally, we apply the lineage-aware Image-space metric (Section~\ref{subsec:image-space-metric})
to a time-series containing multiple division events, demonstrating the feasibility of our approach
in such scenarios.
